Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-06-08 fib2.tex
i think it was a previous typo of mine that got this line commented, or was it on purpose? also trivial image rescaling.
take out all fusion trees and try to explain things using curve diagrams.

\documentclass[aps, prl, letterpaper, twocolumn, superscriptaddress, notitlepage, 10pt]{revtex4-1}
\usepackage{times}

@@

\usepackage{tikz}
\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,positioning,fit}

@@

@@

\newcommand{\e}{\mathrm{e}}

@@

\begin{abstract}
Classically simulating the dynamics of anyonic excitations in two-dimensional quantum systems is likely intractable in general because such dynamics are sufficient to implement 
universal quantum computation. However, processes of interest for the study of quantum 
error correction in anyon systems are typically drawn from a restricted class that displays 
significant structure over a wide range of system parameters.
We exploit this structure to classically simulate, and thereby demonstrate the success of, an 
error-correction protocol for a quantum memory based on the universal Fibonacci anyon 
model.  We numerically simulate a phenomenological model of the system and noise 
processes on lattice sizes of up to 
$128\times128$ sites, and find a lower bound on the error-correction threshold of 
approximately $12.5\%$, which is comparable to those previously known for abelian and 
(non-universal) nonabelian anyon models.
\end{abstract}

@@

In order to efficiently classically simulate an error-correction protocol for a topologically ordered quantum memory, it is necessary to simulate the physical noise processes, the decoding algorithm, and the physical recovery operations. 
Decoding algorithms are typically designed to run efficiently on a classical computer, but there is generally no guarantee that the noise and recovery processes should be classically simulable.
Because of this, almost all of the sizable research effort on active quantum error correction for topological systems has focused on the case of abelian anyons~\cite{Dennis2002, Duclos-Cianci2010, Duclos-Cianci2010a, Wang2010, Wang2010a, Duclos-Cianci2013, Bravyi2011, Bombin2012, Wootton2012, Anwar2014, Watson2014, Hutter2014a, Bravyi2014, Wootton2015, Fowler2015, Andrist2015}.
%~\cite{Dennis2002, Duclos-Cianci2010, Bravyi2011, Bravyi2014}.
Systems of abelian anyons are well suited to studying quantum error correction because (at the RG fixed point) noise and recovery processes can be efficiently simulated numerically, allowing lattice simulations of decoding with over 1 million sites~\cite{Duclos-Cianci2010}. 
Standard algorithms are specifically tailored to exploit the abelian nature of these particles, particularly that abelian anyons cannot be used for quantum computation.

Recent investigations have begun to explore quantum error correction for nonabelian anyon 
models~\cite{Brell2013, Wootton2013, Hutter2014, Wootton2015b}. Nonabelian anyon models are especially interesting 
because braiding and fusion of these anyons in general allows for the implementation of universal quantum 
computation. However, the initial studies of error-correction in nonabelian anyon systems have focused on specific models, such as the Ising 
anyons~\cite{Brell2013} and the so-called $\Phi$-$\Lambda$ 
model~\cite{Wootton2013, Hutter2014} that, while nonabelian, are not universal. The general dynamics of these particular anyon models is known to be efficiently classically simulable, a fact
that was exploited to enable efficient simulation of error correction in these systems. When considering more general anyon models, their ability to perform universal quantum computation would seem a significant barrier to their simulation on a classical computer. While simulation of general dynamics does indeed seem intractable, we argue that the kinds of processes that are typical of thermal noise are sufficiently structured to allow for their classical simulation in the regimes where we expect successful error correction to be possible. This insight allows us to simulate the noise and recovery processes for a quantum code based on a universal anyon model.

@@

%------------------------------------------------------------------------------------------------------------%
\paragraph{Modeling Fibonacci anyons.}

The topological features of a physical anyon model are abstractly described by a unitary 
modular tensor category~\cite{Wang2010b}. This object contains the data describing the 
types of anyonic particles, as well as the results of topological operations such as braiding and 
fusion of these anyons. The Fibonacci anyon model consists of only one non-trivial particle 
type, conventionally labelled $\tau$. Denoting the vacuum by $\mathbb{I}$, we can describe 
the possible fusion outcomes of Fibonacci anyons as $\tau\times\tau=\mathbb{I}+\tau$, 
i.e.~two $\tau$ particles can either fuse to vacuum or to a $\tau$ particle.

Associated with each set of fusion outcomes for a fixed number of $\tau$ particles is a basis 
vector for a Hilbert space, the \emph{fusion} space. For $n$ particles of type $\tau$, the 
dimension of this space grows asymptotically as $\varphi^n$ for $\varphi=\frac{1+\sqrt{5}}{2}$ 
the golden ratio. Additionally, there is a global degeneracy associated with the topology of the 
manifold on which the anyons reside. We will consider systems with the topology of a torus, 
which for the Fibonacci anyons gives rise to a 2-fold degeneracy. For convenience 
(in particular to minimize finite-size effects in our numerical computations~\cite{Brell2013}), 
this global space is the one in which we will encode, and demonstrate protection of, quantum 
information.

The fusion outcome of any particular experiment depends on the history of the particles, in 
particular how they have braided around one another. Braiding operations are treated as 
unitary actions on the fusion space. Pair-creation events should be considered isometries 
that enlarge the fusion space, while fusion events correspond to projective measurements 
of combined charge of some number of particles. For detailed description of these 
processes in Fibonacci anyons, see e.g.~Ref.~\cite{Nayak2008} and references therein. 
Braiding of Fibonacci anyons and measurements of fusion outcomes is known to be 
sufficient to implement universal quantum computation~\cite{Freedman2002, Nayak2008}.

The dynamics and history of a system of anyons can be conveniently represented by a graphical calculus, which can be regarded as a representation of the world-lines of the particles. These diagrams can be manipulated by $R$-moves, corresponding to braiding processes:
	\begin{align}
	\label{e:rmove}
		\begin{tikzpicture}[baseline=0.5cm,scale=0.3]		
			\draw(0,0)--(0,1);
			\draw plot [smooth, tension=1] coordinates { (0,1) (0.75,1.8) (-0.6,2.45) (-1,3)};
			\draw[color=white, line width=7] plot [smooth, tension=1] coordinates { (0,1) (-0.75,1.8) (0.6,2.45) (1,3)};
			\draw plot [smooth, tension=1] coordinates { (0,1) (-0.75,1.8) (0.6,2.45) (1,3)};
			\node at (0,0) [anchor=west] {{\scriptsize $c$}};
			\node at (-1,3) [anchor=east] {{\scriptsize $b$}};
			\node at (1,3) [anchor=west] {{\scriptsize $a$}};
		\end{tikzpicture}
		&\;\;=\;\;R^{ab}_c\cdot\;
		\begin{tikzpicture}[baseline=0.5cm,scale=0.3]		
			\draw(0,0)--(0,1);
			\draw(0,1)--(-1,2)--(-1,3);
			\draw(0,1)--(1,2)--(1,3);
			\node at (0,0) [anchor=west] {{\scriptsize $c$}};
			\node at (-1,3) [anchor=east] {{\scriptsize $b$}};
			\node at (1,3) [anchor=west] {{\scriptsize $a$}};
		\end{tikzpicture}
	\end{align}
where $a$, $b$, and $c$ represent particle types, or $F$-moves, representing the associativity rule for fusion processes, 
	\begin{align}
		\label{e:fmove}
		\begin{tikzpicture}[baseline=0.6cm,scale=0.3]
			\draw(0,0)--(0,1)--(-2,3);
			\draw(0,3)--(-1,2);
			\draw(2,3)--(0,1);
			\node at (-0.45,1.6) [anchor=north east] {{\scriptsize $i$}};
			\node at (0,0) [anchor=west] {{\scriptsize $d$}};
			\node at (2,3) [anchor=east] {{\scriptsize $c$}};
			\node at (0,3) [anchor=east] {{\scriptsize $b$}};
			\node at (-2,3) [anchor=east] {{\scriptsize $a$}};
		\end{tikzpicture}
		&\;\;=\;\;\sum_j[F^{abc}_d]_{ij}\cdot\;
		\begin{tikzpicture}[baseline=0.6cm,scale=0.3]		
			\draw(0,0)--(0,1)--(-2,3);
			\draw(0,3)--(1,2);
			\draw(2,3)--(0,1);
			\node at (0.42,1.63) [anchor=north west] {{\scriptsize $j$}};
			\node at (0,0) [anchor=east] {{\scriptsize $d$}};
			\node at (2,3) [anchor=east] {{\scriptsize $c$}};
			\node at (0,3) [anchor=east] {{\scriptsize $b$}};
			\node at (-2,3) [anchor=east] {{\scriptsize $a$}};
		\end{tikzpicture}
	\end{align}

These manipulations encode the data of the unitary modular tensor category, and can be used to represent the state of a system of anyons as a linear combination of fusion trees. For the Fibonacci anyons, the non-trivial $R$ and $F$ moves are 
\begin{equation*}
	R_{\mathbb{I}}^{\tau\tau} = \e^{\frac{-4\pi i}{5}} 
	\ \ , \ \
	R_\tau^{\tau\tau}= \e^{\frac{3\pi i}{5}} 
	\ \ , \ \
	F_{\tau}^{\tau\tau\tau} = \begin{pmatrix}\varphi^{-1}&\varphi^{-\frac{1}{2}}\\\varphi^{-\frac{1}{2}}&-\varphi^{-1}\end{pmatrix} \,,
\end{equation*}
where the matrix is given in a basis labelled $(\mathbb{I},\tau)$.

We use a phenomenological model of Fibonacci anyon dynamics which neglects any 
microscopic details of the system. 
This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the 
anyon model correspond to large length-scale physics, while the microscopic physics plays 
a less important (and non-universal) role.
Our simulation is in effect that of a $(2+1)$-dimensional topological quantum field theory, which allows for topological processes such as pair-creation, braiding, and fusion of anyonic charges. 
This is modeled on a periodic $L\times L$ square lattice $\Lambda$ of \emph{tiles} on which anyons can reside. 
This treatment applies equally well to lattice models, where anyons can only be located at a discrete set of points, or continuum models, where a discretization of the manifold must be chosen in order to perform charge measurements. 
In the former case, each tile of $\Lambda$ represents a lattice site, while in the latter case each tile represents a measurement region. For more details of an analogous model, see~\cite{Brell2013}.

As was noted in~\cite{Pfeifer2014}, when dealing with two-dimensional 
anyon dynamics, it is not sufficient to consider only
the algebraic (fusion/braiding) properties of the anyons,
one must also account for the algebra of observables. \cggb{This sentence seems odd here. Consider moving?}

@@

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system. We imagine an idealized Hamiltonian for this 
system of the form
\begin{align}
	H=-\sum_{i\in \Lambda}\proj{\mathbb{I}}_i\;,\label{e:hamiltonian}
\end{align}
with $\proj{\mathbb{I}}_i$ the projector to charge $\mathbb{I}$ at site (or tile) $i$, i.e.~the ground 
space of the model has vacuum (or no anyon) at each site. Given the toroidal boundary 
conditions, this space is two-fold degenerate as noted above. We identify this ground space 
as the codespace of our model. Typical realistic noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons on neighboring sites of $\Lambda$. It was 
seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study. 
Such pair-creation events act on neighboring pairs of sites,
and so may be associated with an edge of the lattice, chosen 
uniformly at random at each time step of the simulation. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

In order to perform a logical error on our code, a noise process must have support on a 
ribbon running around a homologically nontrivial loop of the lattice. These correspond to 
processes in which anyonic charge is transported around the non-trivial loop before 
annihilating to vacuum.

Our error-correction algorithm is
based on a hierarchical clustering algorithm~\cite{Hastie2009, Wootton2015b},
and follows a similar strategy to the hard-decision renormalization group decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, producing a \emph{syndrome} of occupied sites. Following this, it forms clusters of nearby anyons and 
measures the total charge within each cluster.
Clusters with trivial charge are discarded, and
the rest are joined (agglomeratively~\cite{Hastie2009})
at linearly increasing length scales. This iterative process concludes when there is at most a single cluster remaining (see \Fref{f:decode}).
Note that unlike the case of abelian anyons, the decoder
cannot determine the charge of each cluster given only the syndrome information (as in \cite{Bravyi2011}),
and so will query (measure) the system to find this.
Our simulation proceeds in this way as a dialogue between decoder
and system, terminating when either the decoder itself terminates or when it performs a homologically nontrivial operation, thereby altering the codespace (and failing to perform successful information recovery).
Note that, as was found in Ref.~\cite{Brell2013}, we expect that our qualitative results may be reproduced by most alternative families of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a clustering decoder is 
its simplicity and flexibility, and the fact that its clustering scheme is compatible with the structure in the noise processes that allows us to classically simulate them.

@@

Although simulating pair creation, braiding, and fusion of $n$ Fibonacci anyons is equivalent 
in computational power to universal quantum computing (and thus unlikely to be classically 
tractable), noise processes and error-correction procedures have structure that we can 
exploit to efficiently simulate typical processes of interest. In particular, those 
processes in which we expect error-correction to succeed are also those that we expect to 
be able to efficiently simulate for the following heuristic reasons (we leave a more rigorous analysis of simulability for noise and error correction processes as an open problem).

Below the (bond) percolation threshold for (say) a 2D square lattice, we expect random sets of 
bonds to decompose into separate connected components 
of average size $O(\log(n))$ and variance $O(1)$~\cite{Bazant2000}.
Each noise process in our model is associated with a (randomly distributed) edge, and so 
disconnected components correspond to sets of anyons that could not have interacted at any 
point in their history. 
We are free to neglect the degrees of freedom associated with braiding between components because each component has trivial total charge.
This allows us to simulate the braiding processes within each component separately. 
In other words: the quantum state in the fusion space of all anyons factorizes into 
a tensor product over components. 
Since each 
component has size only $O(\log(n))$, we can typically simulate these dynamics efficiently 
because the resulting fusion space has dimension $O(\mathrm{poly}(n))$. There are rare 
random processes that violate this reasoning, but processes such as these are suppressed 
exponentially in the lattice size $L$~\cite{Grimmett1989}.

@@

Of course, if we were to consider noise so strong that 
the components percolate over the entire lattice, we would no 
longer expect to be able to efficiently simulate the noise and recovery processes in the system. 
However, recall that logical errors in our system correspond to processes that act over a non-trivial loop of the torus. 
Below the point at which the decoder begins to create such nontrivial loops, our simulation remains efficient,
while it begins to declare failure whenever non-trivial logical 
operations occur. This allows us to simulate large lattice sizes 
in the successful error-correction (sub-percolated) regime, including up to $L=128$.

@@

	\includegraphics[width=1.0\columnwidth]{pic-syndrome.pdf}\\[10pt]
	(e)\ \ 
	\begin{tikzpicture}[baseline=0.7cm,scale=0.3, xscale=0.9]
			\begin{scope}[xshift=2cm,yshift=2cm]		
				\draw(1,0)--(0,1)[dashed];
				\draw(0,1)--(-1,2)--(-1,3);
				\draw(0,1)--(1,2)--(1,3);

>>>>

%\documentclass[aps, prl, letterpaper, twocolumn, superscriptaddress, notitlepage, 10pt]{revtex4-1}
\documentclass[aps, prl, letterpaper, twocolumn, superscriptaddress, notitlepage, 10pt]{revtex4}
\usepackage{times}

@@

%\usepackage{tikz}
%\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,positioning,fit}

@@

\usepackage[normalem]{ulem}

@@

\newcommand{\e}{\mathrm{e}}
\newcommand{\vac}{\mathbb{I}}

@@

\begin{abstract}
Classically simulating the dynamics of anyonic excitations in two-dimensional quantum systems is likely intractable in general because such dynamics are sufficient to implement 
universal quantum computation. However, processes of interest for the study of quantum 
error correction in anyon systems are typically drawn from a restricted class that displays 
significant structure \dude{sparsity?} over a wide range of system parameters.
We exploit this structure \dude{sparsity?} to classically simulate, and thereby demonstrate the success of, an 
error-correction protocol for a quantum memory based on the universal Fibonacci anyon 
model.  We numerically simulate a phenomenological model of the system and noise 
processes on lattice sizes of up to 
$128\times128$ sites, and find a lower bound on the error-correction threshold of 
approximately $12.5\%$, which is comparable to those previously known for abelian and 
(non-universal) nonabelian anyon models.
\end{abstract}

@@

In order to efficiently classically simulate an error-correction protocol for 
a topologically ordered quantum memory, it is necessary to simulate 
the physical noise processes, the decoding algorithm, and the physical recovery operations. 
Decoding algorithms are typically designed to run efficiently on a 
classical computer, but there is generally no guarantee that the 
noise and recovery processes should be classically simulable.
Because of this, almost all of the sizable research effort 
on active quantum error correction for topological systems has focused 
on the case of abelian anyons~\cite{Dennis2002, Duclos-Cianci2010, Duclos-Cianci2010a, Wang2010, Wang2010a, Duclos-Cianci2013, Bravyi2011, Bombin2012, Wootton2012, Anwar2014, Watson2014, Hutter2014a, Bravyi2014, Wootton2015, Fowler2015, Andrist2015}.
Systems of abelian anyons are well suited to studying quantum 
error correction because (at the RG fixed point) noise and 
recovery processes can be efficiently simulated numerically, allowing lattice simulations 
of decoding with over 1 million sites~\cite{Duclos-Cianci2010}. 
Standard algorithms are specifically tailored to exploit the abelian nature 
of these particles, particularly that abelian anyons cannot be used for quantum computation.

Recent investigations have begun to explore quantum error correction for nonabelian anyon 
models~\cite{Brell2013, Wootton2013, Hutter2014, Wootton2015b}. Nonabelian anyon models are especially interesting 
because braiding and fusion of these anyons in general allows for the implementation of universal quantum 
computation. However, the initial studies of error-correction in nonabelian anyon systems have focused on specific models, such as the Ising 
anyons~\cite{Brell2013} and the so-called $\Phi$-$\Lambda$ 
model~\cite{Wootton2013, Hutter2014} that, while nonabelian, are not universal. The general dynamics of these particular anyon models is known to be efficiently classically simulable, a fact
that was exploited to enable efficient simulation of error correction 
in these systems. When considering more general anyon models, their 
ability to perform universal quantum computation would seem a significant 
barrier to their simulation on a classical computer. While simulation 
of general dynamics does indeed seem intractable, we argue that 
the kinds of processes that are typical of thermal noise 
are sufficiently structured  \dude{sparse?} to allow for their classical simulation 
in the regimes where we expect successful error correction to 
be possible. This insight allows us to simulate the noise 
and recovery processes for a quantum code based on a universal anyon model.

@@

%------------------------------------------------------------------------------------------------------------%
%\paragraph{Modeling Fibonacci anyons.}
%\paragraph{Topological quantum field theory.}
\paragraph{Topological model.}

%The topological features of a physical anyon model are abstractly described by a unitary 
%modular tensor category~\cite{Wang2010b}. This object contains the data describing the 
%types of anyonic particles, as well as the results of topological operations such as braiding and 
%fusion of these anyons. The Fibonacci anyon model consists of only one non-trivial particle 
%type, conventionally labelled $\tau$. Denoting the vacuum by $\vac$, we can describe 
%the possible fusion outcomes of Fibonacci anyons as $\tau\times\tau=\vac+\tau$, 
%i.e.~two $\tau$ particles can either fuse to vacuum or to a $\tau$ particle.
%
%Associated with each set of fusion outcomes for a fixed number of $\tau$ particles is a basis 
%vector for a Hilbert space, the \emph{fusion} space. For $n$ particles of type $\tau$, the 
%dimension of this space grows asymptotically as $\varphi^n$ for $\varphi=\frac{1+\sqrt{5}}{2}$ 
%the golden ratio. Additionally, there is a global degeneracy associated with the topology of the 
%manifold on which the anyons reside. We will consider systems with the topology of a torus, 
%which for the Fibonacci anyons gives rise to a 2-fold degeneracy. For convenience 
%(in particular to minimize finite-size effects in our numerical computations~\cite{Brell2013}), 
%this global space is the one in which we will encode, and demonstrate protection of, quantum 
%information.
%
%The fusion outcome of any particular experiment depends on the history of the particles, in 
%particular how they have braided around one another. Braiding operations are treated as 
%unitary actions on the fusion space. Pair-creation events should be considered isometries 
%that enlarge the fusion space, while fusion events correspond to projective measurements 
%of combined charge of some number of particles. For detailed description of these 
%processes in Fibonacci anyons, see e.g.~Ref.~\cite{Nayak2008} and references therein. 
%Braiding of Fibonacci anyons and measurements of fusion outcomes is known to be 
%sufficient to implement universal quantum computation~\cite{Freedman2002, Nayak2008}.

% From Picture TQFTs paper:
%Within physics, TQFTs are referred to as “anyonic systems” [Wil][DFNSS]. These are
%2-dimensional quantum mechanical systems with point like excitations (variously
%called “quas-particle” or just “particle”, anyon, or perhaps “nonabelion”) which
%under exchange exhibit exotic statistics: a nontrival representation of the braid
%groups acting on a finite dimensional Hilbert space V consisting of “internal de-
%grees of freedom”. Since these “internal degrees of freedom” sound mysterious,
%we note that this information is accessed by fusion: fuse pairs of anyons along
%a well defined trajectory and observe the outcome.
%Statistics of anyons is described by unitary representations of the braid groups

We consider a 
2-dimensional system with
point-like excitations or quasiparticles known as anyons.
These anyons show exotic statistics under exchange.
The worldlines of these exchanges form \emph{braids}
in (2+1)-dimensions
which act unitarily on (the Hilbert space of) degrees of freedom known as the \emph{fusion space.}
To access this space we define
the \emph{observables} of the system: these
are regions bounded by circles (Wilson loops)
and measure the total charge of the region.
%The algebraic properties of these processes
These dynamics and observables obey algebraic rules
given by a unitary modular tensor category~\cite{Wang2010b} (UMTC).

%As was noted in~\cite{Pfeifer2014}, when dealing with two-dimensional 
%anyon dynamics, it is not sufficient to consider only
%the algebraic (fusion/braiding) properties of the anyons,
%one must also account for the algebra of observables. \cggb{This sentence seems odd here. Consider moving?}

From the perspective of
\emph{topological quantum field theory}
(TQFT), anyons are modelled
as punctures in a surface~\cite{Pfeifer2014}, and it is not
the anyons themselves that braid, but the space
around the anyons that becomes braided (twisted).
The generators corresponding to the braid
group generators are called half-twists.
This is the perspective we adopt.
In particular, this shows the action of braiding
on arbitrary observables (Wilson loops) as
these occupy the space around the anyons.

The state of a collection of anyons
can be described by the observables of the system.
An important difference between abelian and non-abelian
theories is this: the outcome of measurements in an abelian theory
are determined by particle content only.
That is, the fusion space is one-dimensional (a phase).
This is no longer the case in a non-abelian theory.
An example is Fibonacci: $\tau\times\tau=\tau+\vac$
\emph{picture here?}
%specify a state requires nesting multiple wilson lines.

Non-intersecting loops commute (why?), and so
a basis for the state can be built from a maximal
%set of non-intersecting loops.
set of disjoint and nested loops.
\emph{picture?}
%Observables measure the charge within a
%region bounded by a directed closed loop.
For three anyons, we relate two
possible ways of nesting loops %these measurements
via $F$-moves:
\begin{align}
\label{e:fmove}
\includegraphics[]{pic-fmove.pdf}
\end{align}
Here we have shown particles with charges $a$, $b$, and $c$
as well as total charge $d$.
The lhs shows a state where the anyons with charge $a$ and $b$ are
observed to have total charge $i$.
The rhs involves a superposition (summation)
over charge states $j$ of the anyons with charge $b$ and $c$.
The green line serves to linearly
order the anyons, and track
deformations in the underlying space 
(the history of braiding processes).
It also denotes the direction along
which $F$-moves occur.

%Here we show how braids effect observables:
%Here we show the $R$-move, corresponding to a (clockwise) braiding processes:
A clockwise braiding process is represented by an $R$-move:
\begin{align}
\label{e:rmove}
\includegraphics[]{pic-rmove.pdf}
\end{align}
%where $a$, $b$, and $c$ represent particle types, and $F$-moves,
%representing the associativity rule for fusion processes, 
This is meant to show a \emph{half-twist} of the space
inside the observable with value $c$
holding two anyons (punctures) with charges $a$ and $b$.
We are free to draw loops around charges,
and in general braiding will twist these loops in different
ways.
\emph{picture?}

These pictures (sans quantum) are called 
\emph{curve diagrams} and are used in
the study of braid groups \cite{Dehornoy2002}.
In general, such a curve diagram is 
a continuous embedding of a line that passes through 
each puncture once. 
Remarkably, any two such linear orderings are related by
braiding (half-twists).

The $R$ and $F$-moves obey composition of diagrams.
\emph{picture?}
Fusion is the ``reverse'' of composition: replace the
interior of a disc by a single charge.
\emph{picture?}

%The dynamics and history of a system of anyons can be conveniently represented by a graphical calculus, which can be regarded as a representation of the world-lines of the particles. These diagrams can be manipulated by $R$-moves, corresponding to braiding processes:
%	\begin{align}
%	\label{e:rmove}
%		\begin{tikzpicture}[baseline=0.5cm,scale=0.3]		
%			\draw(0,0)--(0,1);
%			\draw plot [smooth, tension=1] coordinates { (0,1) (0.75,1.8) (-0.6,2.45) (-1,3)};
%			\draw[color=white, line width=7] plot [smooth, tension=1] coordinates { (0,1) (-0.75,1.8) (0.6,2.45) (1,3)};
%			\draw plot [smooth, tension=1] coordinates { (0,1) (-0.75,1.8) (0.6,2.45) (1,3)};
%			\node at (0,0) [anchor=west] {{\scriptsize $c$}};
%			\node at (-1,3) [anchor=east] {{\scriptsize $b$}};
%			\node at (1,3) [anchor=west] {{\scriptsize $a$}};
%		\end{tikzpicture}
%		&\;\;=\;\;R^{ab}_c\cdot\;
%		\begin{tikzpicture}[baseline=0.5cm,scale=0.3]		
%			\draw(0,0)--(0,1);
%			\draw(0,1)--(-1,2)--(-1,3);
%			\draw(0,1)--(1,2)--(1,3);
%			\node at (0,0) [anchor=west] {{\scriptsize $c$}};
%			\node at (-1,3) [anchor=east] {{\scriptsize $b$}};
%			\node at (1,3) [anchor=west] {{\scriptsize $a$}};
%		\end{tikzpicture}
%	\end{align}
%where $a$, $b$, and $c$ represent particle types, or $F$-moves, representing the associativity rule for fusion processes, 
%	\begin{align}
%		\label{e:fmove}
%		\begin{tikzpicture}[baseline=0.6cm,scale=0.3]
%			\draw(0,0)--(0,1)--(-2,3);
%			\draw(0,3)--(-1,2);
%			\draw(2,3)--(0,1);
%			\node at (-0.45,1.6) [anchor=north east] {{\scriptsize $i$}};
%			\node at (0,0) [anchor=west] {{\scriptsize $d$}};
%			\node at (2,3) [anchor=east] {{\scriptsize $c$}};
%			\node at (0,3) [anchor=east] {{\scriptsize $b$}};
%			\node at (-2,3) [anchor=east] {{\scriptsize $a$}};
%		\end{tikzpicture}
%		&\;\;=\;\;\sum_j[F^{abc}_d]_{ij}\cdot\;
%		\begin{tikzpicture}[baseline=0.6cm,scale=0.3]		
%			\draw(0,0)--(0,1)--(-2,3);
%			\draw(0,3)--(1,2);
%			\draw(2,3)--(0,1);
%			\node at (0.42,1.63) [anchor=north west] {{\scriptsize $j$}};
%			\node at (0,0) [anchor=east] {{\scriptsize $d$}};
%			\node at (2,3) [anchor=east] {{\scriptsize $c$}};
%			\node at (0,3) [anchor=east] {{\scriptsize $b$}};
%			\node at (-2,3) [anchor=east] {{\scriptsize $a$}};
%		\end{tikzpicture}
%	\end{align}

%These manipulations encode the data of the unitary modular tensor 
%category, and can be used to represent the state of 
%a system of anyons as a linear combination of fusion trees.

For the Fibonacci anyons, the non-trivial $R$ and $F$ moves are 
\begin{equation*}
	R_{\vac}^{\tau\tau} = \e^{\frac{-4\pi i}{5}} 
	\ \ , \ \
	R_\tau^{\tau\tau}= \e^{\frac{3\pi i}{5}} 
	\ \ , \ \
	F_{\tau}^{\tau\tau\tau} = \begin{pmatrix}\varphi^{-1}&\varphi^{-\frac{1}{2}}\\\varphi^{-\frac{1}{2}}&-\varphi^{-1}\end{pmatrix} \,,
\end{equation*}
where the matrix is given in a basis labelled $(\vac,\tau)$.

%We use a phenomenological model of Fibonacci anyon dynamics which neglects any 
%microscopic details of the system. 
%This is consistent with the principles of topologically 
%ordered systems and anyonic physics, where the key universal features describing the 
%anyon model correspond to large length-scale physics, while the microscopic physics plays 
%a less important (and non-universal) role.
%Our simulation is in effect that of a $(2+1)$-dimensional topological 
%quantum field theory, which allows for topological processes such as 
%pair-creation, braiding, and fusion of anyonic charges. 
%This is modeled on a periodic $L\times L$ square lattice $\Lambda$ of \emph{tiles} on which anyons can reside. 
%This treatment applies equally well to lattice models, where anyons 
%can only be located at a discrete set of points, 
%or continuum models, where a discretization of the manifold must 
%be chosen in order to perform charge measurements. 
%In the former case, each tile of $\Lambda$ represents a 
%lattice site, while in the latter case each tile represents 
%a measurement region. For more details of an analogous model, see~\cite{Brell2013}.

@@

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system. We imagine an idealized Hamiltonian for this 
system of the form
\begin{align}
	H=-\sum_{i\in \Lambda}\proj{\vac}_i\;,\label{e:hamiltonian}
\end{align}
with $\proj{\vac}_i$ the projector to charge $\vac$ at site (or tile) $i$, i.e.~the ground 
space of the model has vacuum (or no anyon) at each site. Given the toroidal boundary 
conditions, this space is two-fold degenerate as noted above. We identify this ground space 
as the codespace of our model. Typical realistic noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons on neighboring sites of $\Lambda$. It was 
seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study. 
Such pair-creation events act on neighboring pairs of sites,
and so may be associated with an edge of the lattice, chosen 
uniformly at random at each time step of the simulation. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

In order to perform a logical error on our code, a noise process must have support on a 
ribbon running around a homologically nontrivial loop of the lattice.
\dude{any homologically nontrivial region is allowed for this, not just ribbons}
These correspond to 
processes in which anyonic charge is transported around the non-trivial loop before 
annihilating to vacuum.

Our error-correction algorithm is
based on a hierarchical clustering algorithm~\cite{Hastie2009, Wootton2015b},
and follows a similar strategy to the hard-decision renormalization group decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, producing a \emph{syndrome} 
of occupied sites. Following this, it forms clusters of nearby anyons and 
measures the total charge within each cluster.
Clusters with trivial charge are discarded, and
the rest are joined (agglomeratively~\cite{Hastie2009})
at linearly increasing length scales. This iterative process concludes when 
there is at most a single cluster remaining (see \Fref{f:decode}).
Note that unlike the case of abelian anyons, the decoder
cannot determine the charge of each cluster given only the syndrome information (as in \cite{Bravyi2011}),
and so will query (measure) the system to find this.
Our simulation proceeds in this way as a dialogue between decoder
and system, terminating when either the decoder itself terminates or 
when it performs a homologically nontrivial operation, thereby altering the 
codespace (and failing to perform successful information recovery).
Note that, as was found in Ref.~\cite{Brell2013}, we expect that 
our qualitative results may be reproduced by most alternative families 
of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a clustering decoder is 
its simplicity and flexibility, and the fact that its clustering 
scheme is compatible with the structure \dude{sparsity?} in the noise 
processes that allows us to classically simulate them.

@@

Although simulating pair creation, braiding, and fusion of $n$ Fibonacci anyons is equivalent 
in computational power to universal quantum computing (and thus unlikely to be classically 
tractable), noise processes and error-correction procedures have structure \dude{sparsity?} that we can 
exploit to efficiently simulate typical processes of interest. In particular, those 
processes in which we expect error-correction to succeed are also those that we expect to 
be able to efficiently simulate for the following heuristic reasons 
(we leave a more rigorous analysis of simulability for noise 
and error correction processes as an open problem).

Below the (bond) percolation threshold for (say) a 2D square lattice, we expect random sets of 
bonds to decompose into separate connected components 
of average size $O(\log(n))$ and variance $O(1)$~\cite{Bazant2000}.
Each noise process in our model is associated with a (randomly distributed) edge, and so 
disconnected components correspond to sets of anyons that could not have interacted at any 
point in their history. 
We are free to neglect the degrees of freedom associated with braiding between components 
because each component has trivial total charge.
This allows us to simulate the braiding processes within each component separately. 
In other words: the quantum state in the fusion space of all anyons factorizes into 
a tensor product over components. 
Since each 
component has size only $O(\log(n))$, we can typically simulate these dynamics efficiently 
because the resulting fusion space has dimension $O(\mathrm{poly}(n))$. There are \sout{rare}
random processes that violate this reasoning, but processes such as these are suppressed 
exponentially in the lattice size $L$~\cite{Grimmett1989}.

@@

Of course, if we were to consider noise so strong that 
the components percolate over the entire lattice, we would no 
longer expect to be able to efficiently simulate the noise and recovery processes in the system. 
However, recall that logical errors in our system correspond to 
processes that act over a non-trivial loop of the torus. 
Below the point at which the decoder begins to create 
such nontrivial loops, our simulation remains efficient \dude{why are we still conflating these},
while it begins to declare failure whenever non-trivial logical 
operations occur. This allows us to simulate large lattice sizes 
in the successful error-correction (sub-percolated) regime, including up to $L=128$.

@@

	\includegraphics[width=0.8\columnwidth]{pic-syndrome.pdf} %\\[10pt]
%	(e)\ \ 
%	\begin{tikzpicture}[baseline=0.7cm,scale=0.3, xscale=0.9]
%			\begin{scope}[xshift=2cm,yshift=2cm]		
%				\draw(1,0)--(0,1)[dashed];
%				\draw(0,1)--(-1,2)--(-1,3);
%				\draw(0,1)--(1,2)--(1,3);
%				\node at (-1,3) [anchor=west] {{\scriptsize $1$}};
